---
title: "P8105_hw2"
author: "Jingwei Ren"
date: "10/2/2018"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)

```
#problem 1
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).
```{r transit_data,message=FALSE}
train_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, `YES` = "TRUE", `NO` = "FALSE")) %>%
  mutate(entry = as.logical(entry))

```


Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The dataset contain variables include:

1.line
2.station name
3.station latitiude
4.station longitude
5.routes serveed
6.entrance type
7.entry indicating whether the station has entry or not
8.vending showing whether the station has vending system
9.ada compliance

Cleaning steps:

I imported the data, and cleaned the column names by the function clean_names in the janitor. Couple variables are selected from the orign dataset and the entry variable is converted to a logical factor. 

The dimension of the resulting dataset is `r nrow(train_data)` rows by `r ncol(train_data)` columns.
```{r}

```
