---
title: "P8105_hw2"
author: "Jingwei Ren"
date: "10/2/2018"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)

```
#problem 1
* Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).
```{r transit_data,message=FALSE}
train_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, `YES` = "TRUE", `NO` = "FALSE")) %>%
  mutate(entry = as.logical(entry))

```


* Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The dataset contain variables include:

1.line

2.station name

3.station latitiude

4.station longitude

5.routes serveed

6.entrance type

7.entry indicating whether the station has entry or not

8.vending showing whether the station has vending system

9.ada compliance

Cleaning steps:

I imported the data, and cleaned the column names by the function clean_names in the janitor. Couple variables are selected from the orign dataset and the entry variable is converted to a logical factor. 

The dimension of the resulting dataset is `r nrow(train_data)` rows by `r ncol(train_data)` columns.
The data is not tidy.


* Answer the following questions using these data:

```{r distinct}
newtrain_data=distinct(train_data, station_name, line, .keep_all = TRUE)
```

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

There are `r nrow(newtrain_data)` distinct stations

How many stations are ADA compliant?

 `r sum(newtrain_data$ada == TRUE)` stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?

```{r}
entrytrue = train_data %>%  
  filter(vending == "NO") %>% 
  filter(entry == "TRUE") %>%
nrow

vendyno = train_data %>%  
  filter(vending == "NO") %>%
  nrow
  

```

`r round(100 * entrytrue/vendyno, digits=2)`% of station entrances / exits without vending allow entrance.

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r reformat_data}
reformat_data = train_data %>%
  gather(key = route_number, value = route_name,route1:route11) 
  
trainA = reformat_data %>%
  filter(route_name == "A")

adatrue = reformat_data %>%
  filter(route_name == "A") %>%
  filter(ada == "TRUE")
```

`r nrow(distinct(trainA,line,station_name))` distinct stations serve the A train.

`r nrow(distinct(adatrue,line, station_name))` distinct stations that serve the A train and are ADA compliant.

#problem2
* Read and clean the Mr. Trash Wheel sheet:

specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)

use reasonable variable names

omit rows that do not include dumpster-specific data

rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer)
```{r clean_sheet1}
trash_wheel = 
  read_excel('data/HealthyHarborWaterWheelTotals2017-9-26.xlsx', sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

* Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).


```{r clean_precipitation_data}
precipitation_2016 = 
  read_excel('data/HealthyHarborWaterWheelTotals2017-9-26.xlsx', sheet = '2016 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = '2016')
precipitation_2017 = 
  read_excel('data/HealthyHarborWaterWheelTotals2017-9-26.xlsx', sheet = '2017 Precipitation', range = 'A2:B14') %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = '2017')

combine_data = full_join(precipitation_2016,precipitation_2017) %>%
  mutate(month = month.name[month])
```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?


